Towards cross-middleware authentication and single sign-on for ARC Grid middleware

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
Some minor changes in tex file

\title{Towards cross-middleware authentication and single sign-on for ARC Grid middleware
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

Dotted line in Figure \ref{fig:HED} shows an alternative path for the information to propagate among MCCs. A service administrator can configure the MCCs according to the interoperability requirements with the other part. For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE’s SoapSender and SoapReceiver)~\cite{WSElink}. Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)~\cite{AShoshani03} service. This shows the flexibility of HED in terms of protocols support.

\begin{figure}
\includegraphics[width=0.75\textwidth]{HED.png}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler – one for the incoming message and one for the outgoing message respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well. In Figure \ref{fig:HED}, the “AuthZ” and “AuthN” sub-modules inside MCCs and Services are examples of SecHandler.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAML2SSO.png}
\caption{SAML2.0 SSO profile in ARC}
\label{fig:SAML2SSO}
\end{figure}
Here is the explanation of steps shown in Figure \ref{fig:SAML2SSO}:
1.The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side. The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is “saml2sp” which is specific for pointing to the SP Service. Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
2.The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself). Then SP Service uses obtained information to compose and issue the SAML authentication request also by utilizing own X.509 certificate (note that in the SAML2.0 SSO profile, X.509 certificates are still needed for IdP and SP) and sends the response back to the user agent.
3.User agent sends the SAML authentication request to the Identity Provider.
4.Identity Provider requires an act of authentication. The authentication mechanism is outside of the SAML2.0 SSO profile. The Shibboleth IdP implementation uses login handlers for authentication. The current user agent implementation is compatible with the Username/Password login handler of Shibboleth IdP. Through the HTTP protocol, the user agent will feed IdP with the username/password which has been given by the caller of user agent interface.
5.Once the authentication has been succeeded, the IdP issues a SAML response including SAML assertion encrypted by destination SP’s public key, and then this SAML response will be delivered by the user agent to the Service Provider.
6.The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context. That assertion includes the AuthnStatement and AttributeStatement elements.
7.The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.
8.The Grid/Web Service checks the AuthnStatement stored in the session context to see if the session is still valid. If valid, service handles the  request and returns the response to the WS client. Note that in the current implementation the service requires WS client to reuse same TCP connection as the one used by user agent is step 5, in order to guarantee  validity of the SSO result.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SLCS.png}
\caption{Short lived credential service}
\label{fig:SLCS}
\end{figure}
The SLCS service is also a Web Service implemented using ARC service interface. The SLCS client is implemented as command-line interface (CLI) utility which implements the user agent and WS client functionalities. The whole process of SLCS invocation is shown in Figure \ref{fig:SLCS} (steps 1‑8). The process resembles one shown in in Figure \ref{fig:SAML2SSO}, except that steps 7 and 8 use SLCS certificate request and response.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation.png}
\caption{Credential delegation service and client}
\label{fig:Delegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation_protocol.png}
\caption{Credential delegation protocol}
\label{fig:Delegation_protocol}
\end{figure}
For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure \ref{fig:Delegation}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure \ref{fig:Delegation}), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegation.png}
\caption{SAML Token delegation service and client}
\label{fig:SAMLDelegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegationMessage.png}
\caption{SOAP message header with SAML delegation token inside}
\label{fig:SAMLDelegationMessage}
\end{figure}
The delegation process is similar to the process in the X.509 credential delegation service, with the following differences:
1.Transport level communication between delegation client and service is not based on mutual authentication between the proxy certificate and the service certificate; instead, it is based on mutual authentication between the service certificates.
2.Delegation service will not generate any certificate request; instead, the delegation client will pick the public key from the peer certificate after a successful SSL/TLS authentication. So step 2 of the delegation protocol is just some confirmation information which informs delegation client to proceed.
3.For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
4.SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.

@@

In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98},{VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way. Gridsite project implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.

>>>>

\title{Towards cross-middleware authentication and single sign-on for ARC Grid middleware
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

Dotted line in Figure \ref{fig:HED} shows an alternative path for the information to propagate among MCCs. A service administrator can configure the MCCs according to the interoperability requirements with the other part. For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE's SoapSender and SoapReceiver)~\cite{WSElink}. Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)~\cite{AShoshani03} service. This shows the flexibility of HED in terms of protocols support.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED.png}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler – one for the incoming message and one for the outgoing message respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well. In Figure \ref{fig:HED}, the "AuthZ" and "AuthN" sub-modules inside MCCs and Services are examples of SecHandler.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO.png}
\caption{SAML2.0 SSO profile in ARC}
\label{fig:SAML2SSO}
\end{figure}
Here is the explanation of steps shown in Figure \ref{fig:SAML2SSO}:
1.The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side. The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is “saml2sp” which is specific for pointing to the SP Service. Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
2.The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself). Then SP Service uses obtained information to compose and issue the SAML authentication request also by utilizing own X.509 certificate (note that in the SAML2.0 SSO profile, X.509 certificates are still needed for IdP and SP) and sends the response back to the user agent.
3.User agent sends the SAML authentication request to the Identity Provider.
4.Identity Provider requires an act of authentication. The authentication mechanism is outside of the SAML2.0 SSO profile. The Shibboleth IdP implementation uses login handlers for authentication. The current user agent implementation is compatible with the Username/Password login handler of Shibboleth IdP. Through the HTTP protocol, the user agent will feed IdP with the username/password which has been given by the caller of user agent interface.
5.Once the authentication has been succeeded, the IdP issues a SAML response including SAML assertion encrypted by destination SP’s public key, and then this SAML response will be delivered by the user agent to the Service Provider.
6.The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context. That assertion includes the AuthnStatement and AttributeStatement elements.
7.The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.
8.The Grid/Web Service checks the AuthnStatement stored in the session context to see if the session is still valid. If valid, service handles the  request and returns the response to the WS client. Note that in the current implementation the service requires WS client to reuse same TCP connection as the one used by user agent is step 5, in order to guarantee  validity of the SSO result.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SLCS.png}
\caption{Short lived credential service}
\label{fig:SLCS}
\end{figure}
The SLCS service is also a Web Service implemented using ARC service interface. The SLCS client is implemented as command-line interface (CLI) utility which implements the user agent and WS client functionalities. The whole process of SLCS invocation is shown in Figure \ref{fig:SLCS} (steps 1‑8). The process resembles one shown in in Figure \ref{fig:SAML2SSO}, except that steps 7 and 8 use SLCS certificate request and response.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{Delegation.png}
\caption{Credential delegation service and client}
\label{fig:Delegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{Delegation_protocol.png}
\caption{Credential delegation protocol}
\label{fig:Delegation_protocol}
\end{figure}
For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure \ref{fig:Delegation}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure \ref{fig:Delegation}), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAMLDelegation.png}
\caption{SAML Token delegation service and client}
\label{fig:SAMLDelegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAMLDelegationMessage.png}
\caption{SOAP message header with SAML delegation token inside}
\label{fig:SAMLDelegationMessage}
\end{figure}
The delegation process is similar to the process in the X.509 credential delegation service, with the following differences:
1.Transport level communication between delegation client and service is not based on mutual authentication between the proxy certificate and the service certificate; instead, it is based on mutual authentication between the service certificates.
2.Delegation service will not generate any certificate request; instead, the delegation client will pick the public key from the peer certificate after a successful SSL/TLS authentication. So step 2 of the delegation protocol is just some confirmation information which informs delegation client to proceed.
3.For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
4.SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.

@@

In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98},~\cite{VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way. Gridsite project implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.
